\section{CHARMM36/EEF1-SB force field energy}

This module is an implementation of the CHARMM36/EEF1-SB force field\cite{doi:10.1021/ct400730n}.
Briefly described, the method is based on the CHARMM36 force field (with CMAP correction), with an implicit solvent model.

Compared to the standard CHARMM36 force field, the backbone torsion angle term is slightly adjusted to give a better balance between alpha-helix and beta-sheets.
Furthermore, charges on ionic side chains and termini are scaled so these appear neutral, and the Coulomb interaction term uses a distance dependent dielectric constant.
The solvent model is a Gaussian-solvent exclusion model.

The recommended ways of using the module are the \texttt{--energy-charmm-cached} or \texttt{--observable-charmm} command line options.

\subsection{Parameters}
Parameters are stored in the \texttt{\$PHAISTOS\_ROOT/modules/charmm/src/energy/parameters} folder in GROMACS's .itp format.
Nuclear charges and atom-types, however, are hard coded in the topology parser at this point.
Physical constants have been set to the values found in the CHARMM program, and all terms are verified to match the CHARMM program to single precision accuracy.
There is an extensive collection of test-cases in the \texttt{\$PHAISTOS\_ROOT/modules/charmm/test} directory which can be used to verify parameters and energies against the CHARMM program.


\subsection{Debug levels}
There are three debug-levels in non-cached terms.
Level 0 gives no debug output. 
Level 1 and higher will display the internal molecular energy for that energy term in kcal and kJ each iteration.
Level 2 and higher will additionally display information about each individual interaction that contributes to the energy of that term.
This can be useful for situations like changing parameters or analyzing interactions in a structure.
The dedicated \texttt{test\_charmm} program can be used to set a debug-level and run single-point energy evaluations.


\subsection{CHARMM36/EEF1-SB angle bend term\\(\texttt{charmm-bond-stretch})}
This term calculates the energy contributions from angle-bend and Urey-Bradley interactions.


\subsection{CHARMM36/EEF1-SB bond stretch term\\(\texttt{charmm-bond-stretch})}
This term calculates the energy contributions from bond stretch interactions between bonded neighbor atoms.

\subsection{CHARMM36/EEF1-SB CMAP correction term\\(\texttt{charmm-cmap})}
This term calculates the CHARMM36/CMAP backbone phi/psi-torsion correction.


\subsection{CHARMM36/EEF1-SB Coulomb term\\(\texttt{charmm-coulomb})}
This term calculates the electrostatic interaction between non-neighboring atoms and "1-4" neighboring atoms. 
This version uses a distance dependent dielectric constant, $\varepsilon = 1.5 \cdot r_{ij}$.
Furthermore, charges on ionic side chains and termini are scaled so these appear neutral, and the Coulomb interaction term uses a distance dependent dielectric constant.


\subsection{CHARMM36/EEF1-SB implicit solvation term\\(\texttt{charmm-implicit-solvent})}
This term calculates the EEF1-SB Gaussian-excluded solvent energy term.
This standard implementation is extremely computationally intensive,
and is primarily intended for debugging purposes.


\subsection{CHARMM36/EEF1-SB torsion angle term\\(\texttt{charmm-torsion})}
This term calculates the energy contribution from torsion angles.


\subsection{CHARMM36/EEF1-SB improper torsion angle term\\(\texttt{charmm-improper-torsion})}
This term calculates the energy contribution from improper torsion angles (out-of-plane bending).


\subsection{CHARMM36/EEF1-SB van der Waals term\\(\texttt{charmm-vdw})}
This term calculates the Lennard-Jones/van der Waals interaction between non-neighboring atoms and "1-4" neighboring atoms. 


\subsection{CHARMM36/EEF1-SB non-bonded\\(\texttt{charmm-non-bonded})}
This term collects the Coulomb, van der Waals, and EEF1-SB implicit solvent energy terms in one term.

\subsection{Cached CHARMM36/EEF1-SB non-bonded\\(\texttt{charmm-non-bonded-cached})}
This term collects the Coulomb, van der Waals, and EEF1-SB implicit solvent energy terms in one more efficient term.
This version is cached, so only interactions that change after a MC move are recalculated.
This is the preferred way of using the CHARMM36/EEF1-SB non-bonded energy during a simulation.

\subsection{Cached CHARMM36/EEF1-SB bonded-term\\(\texttt{charmm-bonded-cached})}

This energy term collects the angle bend, bond stretch, CMAP correction, torsion angle and improper torsion angle terms into one cached term.
This version is cached, so only interactions that change after a MC move are recalculated.
This is the preferred way of using the CHARMM36/EEF1-SB bonded energy during a simulation.
\\\\Since not all bonded terms are degrees of freedom in the move, there are options to ignore evaluation of these terms.
In most MC moves currently available in PHAISTOS (and especially side chain moves), the improper torsion, bond-stretch and bond-angle terms are not sampled, and these can safely be ignored for most purposes.
If these are ignored it is advised to sample backbone angles from the Engh-Huber prior (e.g.~\texttt{--move-crisp-eh}).
Enabling these can sometimes cause large constant energy offsets during the simulation.
This is especially pronounced for the \texttt{--energy-charmm-bond-stretch} term.

\optiontitle{Settings}
\begin{optiontable}
     \option{ignore-bond-angles}{bool}{false}{Ignore bond angle terms.}
     \option{ignore-bond-stretch}{bool}{false}{Ignore bond stretch terms}
     \option{ignore-torsion-angles}{bool}{false}{Ignore torsion angle terms.}
     \option{ignore-improper-torsion-angles}{bool}{false}{Ignore improper torsion angle terms.}
     \option{ignore-cmap-correction}{bool}{false}{Ignore CMAP correction terms.}
\end{optiontable}

